Let's Get Movin': Frontiers of Optimal Transport

### 2018-04-30 reports/paper/paper.tex
./
intro

\subsection{The Optimal Transport Problem and Basic Properties}

@@

@@

\section{Exact Methods}

@@

  \begin{algorithmic}
\Function{GeneticOT}{$\phi, \gamma, k, \b a, \b b, T$}
\State $\mathsf{initial} \gets \b a \b b^T$
\State $\mathsf{family} \gets \mathsf{Array}[\phi]$, initialized to $
\mathsf{initial}$
\State $\mathsf{children} \gets \mathsf{Array}[\phi\gamma]$, initialized to
 empty
  \For{$\mathsf{generation}$ in $1:T$}
    \For{$\mathsf{person}$ in $\mathsf{family, \gamma}$}
      \For{$\mathsf{person}'$ in \Call{Subset}{$\mathsf{family}$}} \Comment
{\textsc{Subset} chooses a subset of size $\gamma$}
\State $\mathsf{child} \gets$ \Call{Marry}{$\mathsf{person}, 
\mathsf{person}'$}
\State $\mathsf{child} \gets$ \Call{Perturb}{$\mathsf{child},k$}
\State Add $\mathsf{child}$ to $\mathsf{children}$
\EndFor
\EndFor
\State Choose the $\phi$ elements in $\mathsf{children}$ with minimal cost for
 \eqref{eq:ot_problem} and set to $\mathsf{family}$
\EndFor
\State\Return{element in $\mathsf{family}$ that achieves minimal cost}
\EndFunction
  \end{algorithmic}
\end{algorithm}
Here \textsc{Marry} picks two probability matrices and averages them
 elementwise, and \textsc{Perturb} chooses random indices $i_1,i_2, j_1,j_2$ and
 moves mass in the probability matrix in such a way that preserves the marginal
 distributions and respects the constraint that $0\le p \le 1$ for all
 components $p$. Such perturbations are repeated $k$ times. $\phi$ is a
 parameter that controls the size of families, and $\gamma$ is a parameter that
 decides the number of offsprings. At each generation, we take the top $\phi$
 matrices that achieve the lowest cost. We hoped that the population drifts
 towards the optimal transport matrix as the number of generation increases.

>>>>

This project should serve as an introductory note and broad overview of both
 theoretical and computational aspects of the
 optimal transport literature. We hope that this note demonstrates both the
 theoretical elegance and practical versatility of optimal transport
 problems, and we hope to bring the reader to the forefront of the
 literature by going through a few recent articles in detail.

The organization of this report is as follows: \cref{sub:problem_intro}
 introduces the
 discrete optimal transport problem and discusses its interpration, dual
 problem, and basic theoretical properties as a linear program. \cref{sec:exact} discusses methods to solve the optimal transport problem \eqref{eq:ot_problem}
 exactly, introducing the Hungarian algorithm \cite{kuhn2010hungarian} in 
\cref{sub:hungarian_algorithm} and presenting a proof of its correctness. The
 Hungarian algorithm is cubic in run-time, which is optimal for exact solution
 of \eqref{eq:ot_problem}, but is often too slow for practical
 applications discussed in \cref{sub:applications}. As such, we discuss
 approximate
 optimization methods in \cref{sec:inexact_methods}, where 
\cref{sub:entropic_regularization_and_sinkhorn_s_algorithm} focuses on the
 Sinkhorn-Knopp algorithm used to solve a convexification of 
\eqref{eq:ot_problem} known as \ref{eq:entropic}. We present an elegant and
 recent upper
 bound of Sinkhorn-Knopp's run-time by
 \cite{altschuler2017near}, as well as numerical experiments and speculations
 about future work. In \cref{sub:genetic_algorithms_}, we attempt to use genetic
 algorithms to approximately solve \eqref{eq:ot_problem}, motivated by recent advances by 
\cite{such2017deep}. In \cref{sec:applications_}, we showcase toy examples of
 using optimal transport in computer vision with the well-known MNIST dataset.
 \cref{sec:conclusion} concludes.

\subsection{The Optimal Transport Problem and Basic Properties}
\label{sub:problem_intro}

@@

Recent advances in machine learning, computer vision, and statistics have
 brought optimal transport problems to significant relevance. Specifically,
 Wasserstein distances have received intense attention from the computer vision
 and statistics communities. For instance, if we consider images as a discrete
 probabilistic distribution over $\mathbb{R}^2$, the Wasserstein distance turns
 out to mimick human perceptions of similarity: Images that are similar to
 humans are close in Wasserstein distances (See \cite{solomon2015convolutional}
 for an illustration of the power of Wasserstein metrics). Other distances over
 distributions, such as the well-known Kullback-Leibler divergence, do not have
 such properties. Such properties of the Wasserstein distance also motivates
 advances in image generation via an incorporation with the increasingly popular
 generative adversarial network (GAN) framework \cite{arjovsky2017wasserstein}.

In terms of statistics, to fit models where likelihood evaluation is costly
 but
 sampling is not (e.g. with data from complicated multistage processes with
 intractible likelihoods), we commonly sample synthetic data from the model, and
 try to match the synthetic data with the observed data by minimizing certain
 distances (known as Approximate Bayesian Computation in the statistics
 literature).
It can be shown that Wasserstein distances should be the distance of choice and
 that Approximate Bayesian Computation via Wasserstein distances has good
 statistical properties \cite{bernton2017inference}.
%Thus, Wasserstein distances are extremely important in computer vision, machine learning, and statistics.

@@

\section{Exact Methods}
\label{sec:exact}

@@

  \begin{algorithmic}
\Function{GeneticOT}{$\phi, \gamma, k, \b a, \b b, T$}
\State $\mathsf{initial} \gets \b a \b b^T$
\State $\mathsf{family} \gets \mathsf{Array}[\phi]$, initialized to $
\mathsf{initial}$
\State $\mathsf{children} \gets \mathsf{Array}[\phi\gamma]$, initialized to
 empty
  \For{$\mathsf{generation}$ in $1:T$}
    \For{$\mathsf{person}$ in $\mathsf{family}$}
      \For{$\mathsf{person}'$ in \Call{Subset}{$\mathsf{family}, \gamma$}} \Comment
{\textsc{Subset} chooses a subset of size $\gamma$}
\State $\mathsf{child} \gets$ \Call{Marry}{$\mathsf{person}, 
\mathsf{person}'$}
\State $\mathsf{child} \gets$ \Call{Perturb}{$\mathsf{child},k$}
\State Add $\mathsf{child}$ to $\mathsf{children}$
\EndFor
\EndFor
\State Choose the $\phi$ elements in $\mathsf{children}$ with minimal cost for
 \eqref{eq:ot_problem} and set to $\mathsf{family}$
\EndFor
\State\Return{element in $\mathsf{family}$ that achieves minimal cost}
\EndFunction
  \end{algorithmic}
\end{algorithm}
Here \textsc{Marry} picks two probability matrices and averages them
 elementwise, and \textsc{Perturb} chooses random indices $i_1,i_2, j_1,j_2$ and
 moves mass in the probability matrix in such a way that preserves the marginal
 distributions and respects the constraint that $0\le p \le 1$ for all
 components $p$. Such perturbations are repeated $k$ times. $\phi$ is a
 parameter that controls the size of families, and $\gamma$ is a parameter that
 decides the number of offsprings. At each generation, we take the top $\phi$
 matrices that achieve the lowest cost. We hoped that the population drifts
 towards the optimal transport matrix as the number of generation increases.
